A user-extensible and adaptable parser architecture

### 2008-08-16 doc/sgai-2008.tex
Small changes to reduce the paper to 14 pages.

Each action contains the code required to deal with a particular type of
log lines, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the actions required from scratch, unless
extending an existing parser.  In the Postfix log parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very skewed distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  It is expected that parsers utilising
this architecture will have a high ratio of rules to actions, hopefully
resulting in simpler rules and clearer distinctions between log lines.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{DELIVERY\_REJECTED}, the
action which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_DATA}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines, but that do not by themselves represent interesting events.  The
third most common action is, perhaps surprisingly, \texttt{UNINTERESTING}:
this action does nothing when executed, allowing uninteresting log lines to
be parsed without causing any effects (it does not imply that the input is
ungrammatical or unparsed).  Generally rules specifying the
\texttt{UNINTERESTING} action are rules to parse log lines which are not
associated with a specific mail, e.g.\ warnings about configuration files
changing.  The framework warns about every unparsed log line, to alert the
user that they need to alter or extend their ruleset; the Postfix log
parser successfully parses every input line in the \numberOFlogFILESall{}
log files it is currently being tested with.  The remaining actions have
only one or two associated rules: in some cases this is because there will
only ever be one type of log line for that action, e.g.\ all log lines
showing that a remote client has connected are matched by a single rule and
handled by the \texttt{CONNECT} action; in other cases the action is
required to address a deficiency in the log files, or a complication which
arises during parsing.  Taking the \texttt{CONNECT} action as an example,
it creates a new data structure for each client connecting, and saves the
data extracted by the rule (in memory, subsequently to be entered into the
database).  If the data structure already exists it is treated as a symptom
of a bug, and the action issues a warning containing the full contents of
the data structure, plus the log line which has just been parsed.

@@

    \item [Line oriented:]  The architecture is line oriented at present:
        there is no facility for rules to consume more input or push unused
        input back onto the input stream (however actions may change input
        lines and cause the modified line to be re-parsed).  This was not a
        deliberate decision, rather a consequence of the line oriented
        nature of the log files; more flexible approaches could be pursued.

@@

Parsing efficiency is an obvious concern when the Postfix log parser
routinely needs to parse very large log files.  The mail server which
generated the log files used in testing the Postfix log parser accepts
approximately 10,000 mails for 700 users a day.  Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  When generating the timing data used in this
section, \numberOFlogFILES{} log files (totaling 10.08 GB,
\numberOFlogLINEShuman{} log lines) were each parsed 10 times and the
parsing times averaged.  Saving results to the database was disabled for
the test runs, as that dominates the run time of the program, and the tests
are aimed at measuring the speed of the Postfix log parser rather than the
speed of the database and the disks the database is stored on.  The
computer used for test runs was a Dell Optiplex 745 described in
\tableref{Computer used to generate statistics}.  Parsing all
\numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}.  The computer was dedicated to
the task of gathering statistics from test runs, and was not used for any
other purpose while test runs were ongoing; any services not necessary for
running the tests were disabled.

@@

\newpage{} % XXX
How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix log parser ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
is \numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the full
ruleset.  A second set of statistics was generated using the minimum
ruleset and the parsing times compared to those generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for optimal, shuffled and reversed orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.  It is clear from \graphref{Percentage parsing time
increase of maximum ruleset over minimum ruleset} that the increased number
of rules has a noticeable performance impact with reverse ordering, and a
lesser impact with shuffled ordering.  The optimal ordering shows a mean
increase of \input{build/include-full-ruleset-vs-minimum-ruleset} in
parsing time for a \numberOFrulesMAXIMUMpercentage{} increase in the number
of rules.  These results show that the architecture scales very well as the
number of rules increases, and that sorting the rules is an important
optimisation enabling this.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}

>>>>

Each action contains the code required to deal with a particular type of
log line, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the actions required from scratch, unless
extending an existing parser.  In the Postfix log parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very skewed distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.  It is expected that parsers
utilising this architecture will have a high ratio of rules to actions,
hopefully resulting in simpler rules and clearer distinctions between
inputs.  \showgraph{build/graph-action-distribution}{Distribution of rules
per action}{Distribution of rules per action} Unsurprisingly the action
with the highest number of associated rules is \texttt{DELIVERY\_REJECTED},
the action which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_DATA}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines, but that do not by themselves represent interesting events.  The
third most common action is, perhaps surprisingly, \texttt{UNINTERESTING}:
this action does nothing when executed, allowing uninteresting log lines to
be parsed without causing any effects (it does not imply that the input is
ungrammatical or unparsed).  Generally rules specifying the
\texttt{UNINTERESTING} action are rules to parse log lines which are not
associated with a specific mail, e.g.\ warnings about configuration files
changing.  The framework warns about every unparsed log line, to alert the
user that they need to alter or extend their ruleset; the Postfix log
parser successfully parses every input line in the \numberOFlogFILESall{}
log files it is currently being tested with.  The remaining actions have
only one or two associated rules: in some cases this is because there will
only ever be one type of log line for that action, e.g.\ all log lines
showing that a remote client has connected are matched by a single rule and
handled by the \texttt{CONNECT} action; in other cases the action is
required to address a deficiency in the log files, or a complication which
arises during parsing.  Taking the \texttt{CONNECT} action as an example,
it creates a new data structure for each client connecting, and saves the
data extracted by the rule (in memory, subsequently to be entered into the
database).  If the data structure already exists it is treated as a symptom
of a bug, and the action issues a warning containing the full contents of
the data structure, plus the log line which has just been parsed.

@@

    \item [Line oriented:]  The architecture is line oriented at present:
        there is no facility for rules to consume more input or push unused
        input back onto the input stream.  This was not a deliberate
        decision, rather a consequence of the line oriented nature of the
        log files; more flexible approaches could be pursued.

@@

Parsing efficiency is an obvious concern when the Postfix log parser
routinely needs to parse very large log files.  The mail server which
generated the log files used in testing the Postfix log parser accepts
approximately 10,000 mails for 700 users a day.  Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  When generating the timing data used in this
section, \numberOFlogFILES{} log files (totaling 10.08 GB,
\numberOFlogLINEShuman{} log lines) were each parsed 10 times and the
parsing times averaged.  Saving results to the database was disabled for
the test runs, as the tests are aimed at measuring the speed of the Postfix
log parser rather than the speed of the database and the disks the database
is stored on.  The computer used for test runs was a Dell Optiplex 745
described in \tableref{Computer used to generate statistics}.  Parsing all
\numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}.  The computer was dedicated to
the task of gathering statistics from test runs, and was not used for any
other purpose while test runs were ongoing; any services not necessary for
running the tests were disabled.

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix log parser ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
is \numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the full
ruleset.  A second set of statistics was generated using the minimum
ruleset and the parsing times compared to those generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for optimal, shuffled and reversed orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.  It is clear from \graphref{Percentage parsing time
increase of maximum ruleset over minimum ruleset} that the increased number
of rules has a noticeable performance impact with reverse ordering, and a
lesser impact with shuffled ordering.  The optimal ordering shows a mean
increase of \input{build/include-full-ruleset-vs-minimum-ruleset} in
parsing time for a \numberOFrulesMAXIMUMpercentage{} increase in the number
of rules.  These results show that the architecture scales very well as the
number of rules increases, and that sorting the rules is an important
optimisation enabling this.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}
